---
output:
  github_document:
    toc: true
    toc_depth: 2
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
knitr::opts_knit$set(root.dir = normalizePath(rprojroot::find_rstudio_root_file()))
```
# demography module documentation

The demography module provides event functions for simutenouesly simulating the
dynamics of evolving urban population and their households at the individual-based level. 
These events are responsible for household formation (marriage, cohabitation, 
leaving parental home), household dissolutions (separation, divorce, break up), 
household expansion (birth).

# Release note

## Version 0.1.0

First release!

# Requirements

## R packages

R packages: dymiumCore, data.table

# Events

## Aging

### Usage 

```{r, eval = FALSE}
event_demography_age <- modules::use("modules/demography/age.R")
event_demography_age$run(world, model = NULL, target = NULL, time_steps = NULL)
```

### Params

* __object__: a World object.
* __model__: NULL, this event doesn't require this argument.
* __target__: NULL, this event doesn't require this argument.
* __time_steps__: a integer vector that contains the time steps in which this event
should be run.

### Description

Increase the age of all individual agents by 1 which is equipvalent to one year. 

### Note

If there are any attributes that should be updated depending on the age of agent
then it can be implemented inside this event. For example, currently there are 
three rules that are active. These rules update marital status, education,
and labour force status of the agent once it has reached some certain age thresholds.
All agents aged 16 and above can be in a relationship (a marriage or a cohabitation)
hence once they turn 16 their marital status will be changed from "not applicable"
to "never married" which allows them to be considered when the marriage event and
the cohabitation event are being simulated.

### Example

```{r}
create_toy_world()

world %>%
  event_demography_age$run(.)
```

## Birth

### Usage 

```{r, eval = FALSE}
event_demography_birth <- modules::use("modules/demography/birth.R")
event_demography_birth$run(world, model = NULL, target = NULL, time_steps = NULL)
```

### Params

* __object__: a [World] object.
* __model__: a named list that contains path to a MATSim config file.
  * __fertility__: a binary model to determine weather the female in question is to give birth or not.
  * __birth_multiplicity__: chance of giving birth to more than one baby. Note that, the current code is only allow for twins but you may modify this to accommodate cases of triplets and more.
  * __birth_sex_ratio__: the chance of giving birth to a female baby vs a male baby.

```{r eval = FALSE}
model <- list(fertility = list(yes = 0.05, no = 0.95),
              birth_multiplicity = list("single" = 0.97, "twins" = 0.03),
              birth_sex_ratio = list(male = 0.51, female = 0.49))
```

* target: default as `NULL` or it can be a named list which determines the number of individual agents to under go the fertility event. For example, if a list `list(yes = 100)` it will garantee that there are 100 individual agents that will give birth. 
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

```{r}
NULL
```

## Death

### Usage 

### Params

* __world__: a World object.
* __model__: a named list.
  * __death__: a binary model.
  
```{r eval = FALSE}
model <- list(death = list(yes = 0.03, no = 0.97))
```

* __target__: default as `NULL` or a named list.
* __time_steps__: a integer vector that contains the time steps in which this event
should be run.

### Description

Simulate deaths of individual agents. It removes the dying individual agents from
the individual database and also updates attributes (such as marital status, household size)
of any agents that are related to the dying agents.

### Note

The marital status of those individual agents whom their partner has died will be
labelled as "widowed". To retrive the data of all agents that have been removed you
through the death event use the `get_removed_data()` method.

### Example

```{r}
NULL
```
 
## Marriage

### Usage 

### Params

* __object__: a [World] object.
* __model__: a named list.
  * __marriage_cohab_male__: a binary model that determines whether a cohabitating couple will get married. This is based on attributes of the male partner.
  * __marriage_no_cohab_male__: a binary model that determines the chance for eligible males to enter the marital market.
  * __marriage_no_cohab_female__: a binary model that determines the chance for eligible females to enter the marital market.
  * __husbandAgeRuleToCreateNewHousehold__: a binary model that determines the chance of the couple to form a new household or merge their households. 

```{r eval = FALSE}
model <- list(
  marriage_cohab_male = list(yes = 0.1, no = 0.9),
  marriage_no_cohab_male = list(yes = 0.1, no = 0.9),
  marriage_no_cohab_female = list(yes = 0.1, no = 0.9),
  husbandAgeRuleToCreateNewHousehold = list(age = 30)
)
```

* __target__: a named list

```{r}
target <- list(
  yes = 100
)
```

* __time_steps__: a integer vector that contains the time steps in which this event
should be run.

### Description

This event forms a marriage relationship between a cohabiting couple and two individuals that are not in a relationship. After the marriage is formed, the newly wedded couple then decide whether to form a new household or to merge their households into one. When forming a new household, regardless of their household formation decision, any related individuals (i.e. dependent children) to both individuals will also follow to the new household.

### Note

As you can see there are four models that shall be supplied to determine the likelihood in each stage of the marriage event. The first stage is to marry cohabiting couples based on the likelihood produces by the `marriage_cohab_male` model. The second stage is to marry eligible, single individual agents. The probabilities for the individual agents to enter the marital market come from the Monte Carlo simulation result performed using the `marriage_no_cohab_male` and `marriage_no_cohab_female` models on the individual agents' attributes. Then all participating individual agents are paired based on a given rule. The rule can be as simple as all agents prefer to match with an agent that has the minimum age difference to theirs. See `StochasticMarriageMarket` and `OptimalMarriageMarket` for the matching strategies available in `modules/demography/marriage.R`. Note that, if there are more agents of one gender than other in the marital market then the number of maximum matches will be equal to the number of individual agents with the lesser number. Those who are not matched will remain single after the event has ended. The current implementation doesn't include marriages between same-sex couples. After the matching step, all newly formed couples will decide whether they will form a new household (both agents leave their current households) or for the wife and her related individuals to join her husband's household. The current implementation applies a very simple rule which is likely to be wrong and should be replaced if there is a better assumption or model. The current rule is that for all newly wedded couples if their male partner aged more than the number given in `husbandAgeRuleToCreateNewHousehold` they will create a new household.

TLDR;

- Simulate marriages for comhabiting couples and single individuals.
- Aged 16 and above to be able to marry.
- Doesn't consider marriages between same-sex couples.
- The matching strategy for pairing individuals can be configured, either optimal or stochastic.

### Example

```{r}
NULL
```

## Separation

### Usage 

### Params

* __object__: a World object.
* __model__: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* __target__: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

Separation is the first step before couples can be officially divorced. This module assume that no couples will recoupled once they have decided to separate. 

## Divorce


### Usage 

### Params

* __object__: a World object.
* __model__: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* __target__: NULL
* __time_steps__: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

When divorce is triggered for a separted individual, his/her ex-partner will also under go divorce **if** the marital status of his/her ex-partner is still 'separated'.  

## Cohabitation

### Usage 

### Params

* object: a [World] object.
* model: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

## Breakup

### Usage 

### Params

* object: a [World] object.
* model: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

## Leave parental home

### Usage 

### Params

* object: a [World] object.
* model: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

## Migration

### Usage 

### Params

* object: a [World] object.
* model: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

All operations are done within your R environment with no external dependencies. To use any event function please consider using the following commands instead of using `source("path/to/event/script")`.

```{r}
event_demography_age <- modules::use("modules/demography/age.R")
```

# Known issues