---
output:
  github_document:
    toc: true
    toc_depth: 2
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
knitr::opts_knit$set(root.dir = normalizePath(rprojroot::find_rstudio_root_file()))
```
# demography module documentation

The demography module provides event functions for simutenouesly simulating the
dynamics of evolving urban population and their households at the individual-based level. 
These events are responsible for household formation (marriage, cohabitation, 
leaving parental home), household dissolutions (separation, divorce, break up), 
household expansion (birth).

# Release note

## Version 0.1.0

First release!

# Requirements

## R packages

R packages: dymiumCore, data.table

# Events

## Aging

### Usage 

```{r, eval = FALSE}
event_demography_age <- modules::use("modules/demography/age.R")
event_demography_age$run(world, model = NULL, target = NULL, time_steps = NULL)
```

### Params

* object: a [World] object.
* model: NULL, this event doesn't require this argument.
* target: NULL, this event doesn't require this argument.
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

Increase the age of all individual agents by 1 which is equipvalent to one year. 

### Note

If there are any attributes that should be updated depending on the age of agent
then it can be implemented inside this event. For example, currently there are 
three rules that are active. These rules update marital status, education,
and labour force status of the agent once it has reached some certain age thresholds.
All agents aged 16 and above can be in a relationship (a marriage or a cohabitation)
hence once they turn 16 their marital status will be changed from "not applicable"
to "never married" which allows them to be considered when the marriage event and
the cohabitation event are being simulated.

### Example

```{r}
create_toy_world()

world %>%
  event_demography_age$run(.)
```

## Birth

### Usage 

### Params

* object: a [World] object.
* model: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

For female individual agents.


## Death

### Usage 

### Params

* world: a World object.
* model: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example


## Marriage

### Usage 

### Params

* object: a [World] object.
* model: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

Marriages can be divided into two types: marriage with priorly cohabited and marriage without cohabitation. The first type only applies to cohabiting couples who want to get married. The second type of marriage can occur to any non-cohabiting individuals.

### Required Models

- marriage_cohab_male: a binary choice model
- marriage_no_cohab_male: a binary choice model
- marriage_no_cohab_female: a binary choice model

## Separation

### Usage 

### Params

* object: a [World] object.
* model: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

Separation is the first step before couples can be officially divorced. This module assume that no couples will recoupled once they have decided to separate. 

## Divorce


### Usage 

### Params

* object: a [World] object.
* model: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

When divorce is triggered for a separted individual, his/her ex-partner will also under go divorce **if** the marital status of his/her ex-partner is still 'separated'.  

## Cohabitation

### Usage 

### Params

* object: a [World] object.
* model: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

## Breakup

### Usage 

### Params

* object: a [World] object.
* model: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

## Leave parental home

### Usage 

### Params

* object: a [World] object.
* model: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

## Migration

### Usage 

### Params

* object: a [World] object.
* model: a named list that contains path to a MATSim config file.

```{r eval = FALSE}
# config: path to a matsim config file
# lastIteration: a numeric value that denotes the number of iterations for matsim to run
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

### Note

### Example

All operations are done within your R environment with no external dependencies. To use any event function please consider using the following commands instead of using `source("path/to/event/script")`.

```{r}
event_demography_age <- modules::use("modules/demography/age.R")
```

# Known issues