---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# dymiumModules

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/dymium-org/dymiumModules.svg?branch=master)](https://travis-ci.org/dymium-org/dymiumModules)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/dymium-org/dymiumModules?branch=master&svg=true)](https://ci.appveyor.com/project/dymium-org/dymiumModules)
<!-- badges: end -->

This repository contains an R package called `dymiumModules`. The package contains utility functions that facilitate module creation and downloading of dymium compatible modules from any GitHub repository, default to this repo.

To share your module in this repository, fork this repository and send in your module as a pull request. Note that all your module scripts should live under the 'modules' folder. 

# Installation

You can install the latest version of dymiumModules from GitHub with:

```{r eval = FALSE}
remotes::install_github("dymium-org/dymiumModules")
```

Note that, installing this package doesn't give you the modules in this repository. You must use `download_module()` to download them.

# Module directory

- [demography](https://github.com/dymium-org/dymiumModules/tree/master/modules/demography): demographic processes for simulating population and household dynamics.
- [matsim](https://github.com/dymium-org/dymiumModules/tree/master/modules/matsim): an agent-based traffic simulator.
- [test](https://github.com/dymium-org/dymiumModules/tree/master/modules/test): regression testing for dymiumCore.

# Create a new module

__dymiumCore__ provides many helper functions to help you create your first module.
The `use_module(name = "mymodule")` function creates a folder called `modules` 
under your active R project directory. Inside the `modules` folder you will find
another folder called `mymodule` which is the name of your new module and inside
that you will find `logger.R`, `helpers.R`, `constants.R` and a `testthat` folder.

```{r, eval = FALSE}
library("dymiumCore")
use_module(name = "mymodule")
```

To add new events to your newly create module you can use the `use_event()` function
as in the example below.

```{r, eval = FALSE}
use_event("event1", "mymodule")
use_event("event2", "mymodule", with_comments = TRUE) # with recommendations
use_event("event3", "mymodule", with_comments = FALSE) # without recommendations
```

# Download a module from this repository using R

To download a module it is recommended that you use `dymiumCore::download_module()`. 

Before downloading a module, make sure you setup and use an RStudio project. The 
download module will be saved to the 'modules' folder (this will be created if it 
is not already existed) to the root of your working directory. 

```{r, eval = FALSE}
download_module(name = 'test')
```

The `name` argument is the name of the module in this repository that you would like to
download. You may try to download the 'test' module which contains dummy events. 
There are other arguments that you can specify in `download_module()`, please see 
the documentation by calling `?download_module`.
