---
output:
  github_document:
    toc: true
    toc_depth: 2
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  eval = FALSE
)
knitr::opts_knit$set(root.dir = normalizePath(rprojroot::find_rstudio_root_file()))
```

# MATSim module documentation

This module provides the following functionalities:

- `runControl.R`: run MATSim using a MATSim config file.
- `createVISTADemand.R`: assign daily activity patterns to a population and write it as a MATSim plan_v4 file.
- `helpers.R`: reader and edit matsim's config, create a MATSim network_v2 file from shpfiles, assign random geographical locations to activities. 

# Release note

## Version 0.1.0

- Allows basic calls to MATSim APIs from R. 
- Adds functions for preparing MATSim inputs.

# Requiements

## R packages

- R CRAN packages: data.table, modules, StatMatch, checkmate, here, rJava, xml2, sf, glue, furrr, fs
- R github packages: dymium-org/dymiumCore, 

## MATSim

- You will need to install Java. It is recommended to use Java 8 to run MATSim (see related issues here: https://github.com/matsim-org/matsim-code-examples/issues/176, https://github.com/matsim-org/matsim-code-examples/issues/16). 

- The matsim module doesn't shipped with a MATSim java executable file that is required to make this module works. The user must download MATSim from [MATSim website](https://www.matsim.org/downloads/). I recommend to use the latest stable release (version 0.10.0 as of `r Sys.Date()`) with this module. Once, you have downloaded MATSim you must extract the matsim zip file into your matsim module folder (usually under the root folder of your active R project `modules/matsim`). Then rename the extracted file as 'matsim'. Then you are good to go! :) 

# Events

## Run controler

### Usage

```{r, eval = FALSE}
event_matsim_runcontroler <- modules::use("modules/matsim/runControler.R")
event_matsim_runcontroler$run(world, model = NULL, target = NULL, time_steps = NULL)
```

### Parameters

* __world__: a World object.
* __model__: a named list that contains path to a MATSim config file.
  - _config_: a path to a matsim config file
  - _lastIteration_: a numeric value that denotes the number of iterations for matsim to run

```{r eval = FALSE}
model <- list(config = "path/to/config.xml",
              lastIteration = 10)
```

* __target__: NULL
* __time_steps__: a integer vector that contains the time steps in which this event
should be run.


### Description

This script contains the main function which is `event_matsim_runcontroler$run(object, model, target, time_steps)`. The function calls a MATSim configuration file that contains all the settings necessary to run a MATSim scenario. By default, dymium modifies the output directory where MATSim results will be saved to `/<path-to-your-dymium-RStudio-project>/output/matsim`. This is done by creating a modified version (with a `-dymium` suffix) of the supplied config file and saved inside the same directory.

To modify a MATSim config at runtime you may use the suppiled MatsimConfig class in `helpers.R`. This class allows you to modify values inside a MATSim config file using its public methods.

Some MATSim settings can be modified in this script such as the path to a MATSim.jar executable file and the maximum amount of memory for MATSim to utilise. They are stored as `.matsim_setting`. To modify these options open the script and find the lines that are similar to the code in the chuck below.

### Note

Some settings need to be specified such as the maxmimum memory MATSim is allowed to use and the path to the MATSim java executable file. These settings can be changed inside the event script that can be found at `modules/matsim/runControler.R`.

```{r eval = FALSE}
.matsim_setting <- 
  list(
    max_memory = "-Xmx8048m", # maximum amount of memory
    path_to_matsim_jar = here::here('modules/matsim/matsim/matsim-0.11.0-SNAPSHOT.jar') # path to your MATSim.jar
  )
```

### Example

The following example runs an example MATSim scenario for five iterations.

```{r eval = FALSE}
event_matsim_runcontroler <-
  modules::use(here::here('modules/matsim/runControler.R'))

create_toy_world()

for (i in 1:10) {
  world$start_iter(time_step = i, unit = 'year') %>%
    event_matsim_runcontroler$run(model = list(
      config = here::here("modules/matsim/matsim/examples/equil/config.xml"),
      lastIteration = 5
    ))
}
```


## Create a MATSim plan

### Usage

```{r}
event_matsim_createVISTADemand <- modules::use('modules/matsim/createVISTADemand.R')
event_matsim_createVISTADemand$run(world, model = NULL, target = NULL, time_steps = NULL)
```


### Parameters

* world: a World object.
* model: a named list.
  - _vista_persons_: a data.frame that contains a vista person file 
  - _vista_trips_: a data.frame that contains a vista trip file

```{r eval = FALSE}
model <- list(vista_persons = data.frame(),
              vista_trips = data.frame())
```

* target: NULL
* time_steps: a integer vector that contains the time steps in which this event
should be run.

### Description

This event function assigns daily travel activity patterns to Dymium individuals using a statistical matching technique. It uses a travel survey in this case the 2009 VISTA household travel survey of Victoria state, Australia ([link](https://transport.vic.gov.au/about/data-and-research/vista/vista-data-and-publications)). It then parses the fused travel activities to an xml file that follows the format of [MATSim's population_V6](http://www.matsim.org/files/dtd/population_v6.dtd) and save it to your matsim input folder.

### Note

The implementation of this function is very specific to the 2009 VISTA data hence it's most likely not going to work if you replace the vista data with other data. However, this can be used as a template for implementing other travel surveys or can be replaced this step entirely with a proper activity-based model if available.

# Helper functions

- MatsimConfig: an R6 class that can modify a MATSim config file
- create_population_v6: create a population xml file that complies with MATSim's Population V6.
- assign_location: assign a random locations within a shapefile
- shp2xml: convert a network shapfile to a MATSim compatible xml network file

See the `helpers.R` script for more details.

Additionally, there are some functions for binding MATSim's histogram plots into a single animated histogram by iteration number in `plots.R`

# Known issues

## MATSim
The MATSim melbourne scenario has the following unresolved issues

- strategy setting 'SubtourModeChoice' doesn't work. If this setting is used, the error will be raised about modeParam 'pt' and 'walk' not being valid.

## Setup Java 

As of `r Sys.Date()`, Java JDK 11.0.1 must be installed in order for the `rJava` pacakge to work. This issue has been discussed [here](https://github.com/rstudio/rstudio/issues/2254). This is only required if you would like to run the MATSim module using rJava as implemented in `.run_matsim` function inside `runControler.R`. Otherwise, you can use `.run_matsim2` which basically invoke your `matsim.jar` from the OS command, this option requires you to make sure that Java8+ is installed and is callable using the OS command. 

To install the required version of Java click on this [link](https://www.oracle.com/technetwork/java/javase/downloads/java-archive-javase11-5116896.html) and download **Java SE Development Kit 11.0.1** that works on your machine.

You may need to run `R CMD javareconf` on terminal to make sure R detects your new java installation. Make sure that all R sessions are close before running this command.


## rJava failed to load! Ughhhhh!


If you get the follow error from

```r
> library(rJava)
# Error: package or namespace load failed for 'rJava':
#  .onLoad failed in loadNamespace() for 'rJava', details:
#   call: dyn.load(file, DLLpath = DLLpath, ...)
#   error: unable to load shared object '/Users/kevin/Library/R/3.4/library/rJava/libs/rJava.so':
#   dlopen(/Users/kevin/Library/R/3.4/library/rJava/libs/rJava.so, 6): Library not loaded: @rpath/libjvm.dylib
#   Referenced from: /Users/kevin/Library/R/3.4/library/rJava/libs/rJava.so
#   Reason: image not found
```

It may be a good idea to run the following command in your Terminal.

```
R CMD javareconf
```

Then quit RStudio and launch it again to try `library(rJava)`. If the error still persist then Google search is your best friend!